---
title: "5.Figures"
author: "Francis Leblanc"
date: "23/06/2022"
output:   
  html_document:
    toc: yes
    toc_float: yes
    number_sections: false
    code_folding: show 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = FALSE
)
library(Seurat)
library(Signac)
library(tidyverse)
library(ggpubr)
library(cowplot)
library(patchwork)
library(GenomicRanges)
library(ROCR)
library(UpSetR)
library(grid)

theme_set(theme_cowplot())

set.seed(2022)
```


# Fig. 2 - Compare links Z-scores before and after downsampling

```{r}
# pbmc <- readRDS("../data/pbmc_multiome_processed.rds")
links.downsampled <- readRDS("../results/pbmc500_links_r0.01.rds")
links0.1 <- readRDS("../results/links_signac_200null.r0.1.multimode.rds")
ct.count <- readRDS("../data/ct.count.rds")
top.ct <- names(ct.count[ct.count > 50])
```

## match downsampled and full dataset peaks by overlap

```{r}
ol <- findOverlaps(StringToGRanges(links.downsampled$peak),
                   StringToGRanges(links0.1$peak))

colnames(links.downsampled) <- paste0("pbmc500_", colnames(links.downsampled))
combined.res <- cbind(links0.1[ol@to,],
                      as.data.frame(links.downsampled[ol@from,]))

combined.res <- combined.res %>% dplyr::filter(gene == pbmc500_gene)
```

## T-test for Z-scores before and after downsampling
```{r}
# levels by median Z-scores
lvs.by.median <- combined.res %>% 
  group_by(ct_top_auc) %>% 
  summarise(median = median(abs(zscore))) %>% 
  arrange(median) %>% 
  pull(ct_top_auc)

combined.res$ct_top_auc <- factor(combined.res$ct_top_auc, levels = lvs.by.median)

# run t-test by celltype and keep confidence intervals
l.ttests <- lapply(lvs.by.median, function(x){
  df <- combined.res %>% 
    dplyr::filter(ct_top_auc == x) %>% 
    dplyr::select(pbmc500_zscore, zscore) %>% 
    reshape2::melt()
  df$value <- abs(df$value)
  S <- t.test(paired = F, value ~ variable, data = df)
  S <- c(S$conf.int[1:2], S$estimate, S$p.value)
  S
})
df.ttest <- do.call(rbind, l.ttests) %>% as.data.frame()
df.ttest$mean <- df.ttest[,3] - df.ttest[,4]
colnames(df.ttest)[c(1,2,5)] <- c("lower", "upper", "p")
df.ttest$ct <- factor(lvs.by.median, levels = lvs.by.median)
df.ttest <- df.ttest %>% rstatix::add_significance("p")

p.se <-ggplot(df.ttest, aes(x = ct, y = mean, ymin = lower, ymax = upper)) + 
  geom_pointrange() + 
  geom_hline(yintercept=0, linetype="longdash", color = "darkred") +
  annotate("text", y = -1.3, x = df.ttest$ct, label = df.ttest$p.signif) +
  coord_flip() +
  theme(axis.text.y = element_blank()) +
  xlab("") +
  ylab("Differences (mean and standard error) \nbetween peak-gene Z-scores calculated \nafter and before down-sampling MP")

# show cell count in each celltype with > 50 cells
df <- as.data.frame(ct.count)
colnames(df) <- c("ct","freq")

df <- df %>% dplyr::filter(ct %in% top.ct)
df$ct <- factor(df$ct, levels = lvs.by.median)
df[18,] <- c("not.specific", 0) # not.specific peaks

p.bar <- ggplot(df ,aes(x = ct, y = as.numeric(freq))) + 
  geom_col() + 
  ggpubr::rotate() +
  theme(axis.text.y = element_blank()) +
  xlab("") + ylab("Cell count") + NoLegend()

m <- reshape2::melt(combined.res[, c("pbmc500_zscore" ,"zscore", "ct_top_auc")], 
               measure.vars = , 
               id.vars = "ct_top_auc")

### boxplot of Z-scores by celltype before and after downsampling
p.box <- ggplot(m, aes(x=abs(value), y=ct_top_auc, fill=variable)) + 
  geom_point(shape = ".", color="black", size=0.5, alpha=0.5, position = position_jitterdodge()) +
  geom_boxplot(outlier.shape = NA, alpha = 0.5)+
  theme(legend.position = c(0.55, 0.3)) +
  ylab("") + xlab("|Z-score|")+
  scale_fill_manual(values = c("darkred", "gray"), 
                    labels = c("Down-sampled MP" ,"All PBMC")) + 
  labs(fill = "")

ggsave(plot = p.box + p.bar + p.se ,"../figs/5_out/split_boxplot_Zscores_celltypes_links_abs0.1_SE_unpaired.png", width = 14, height = 7)
```

# Fig.S3 - Correlations with number of cells within a celltype

## correlation bimodal peak count vs cell count

```{r}
# get % of links with bimodal null
bimod.count <- links0.1 %>% 
  group_by(ct_top_auc, .drop = FALSE) %>% 
  tally(multimodal) 

bimod.count$ct_marker.count <- links0.1 %>% 
  group_by(ct_top_auc) %>% 
  tally() %>% 
  pull(n)

bimod.count$bimod.pct <- round((100 * bimod.count$n) / bimod.count$ct_marker.count)
bimod.count <- bimod.count %>% 
  arrange(desc(bimod.pct)) %>%
  mutate(ct_top_auc = factor(ct_top_auc, levels = ct_top_auc))

# sort by median Z-scores
df <- as.data.frame(ct.count)
colnames(df) <- c("ct","freq")
df <- df %>% dplyr::filter(ct %in% top.ct)
df$ct <- factor(df$ct, levels = lvs.by.median)
df$bimod.pct <- bimod.count$bimod.pct[match(df$ct, bimod.count$ct_top_auc)]

p <- ggscatter(df, x="freq", y="bimod.pct",
               cor.coef = T, 
               cor.method = "spearman",
               xlab = "Number of cells per cell-type", 
               ylab = "Percentage of bimodal distributions",
               label = "ct",
               repel = T, cor.coef.coord = c(100,60))

ggsave(plot = p, "../figs/5_out/bimodality.pct.vs.cellcount.png", width = 7, height = 5)
```

## Scatter: Relation of celltype count vs median -Log10(P)

```{r}
median_log10P <- links0.1 %>% 
  dplyr::group_by(ct_top_auc) %>% 
  summarise(median.log10p = median(-log10(pvalue))) %>% 
  as.data.frame()

median_log10P$n_cells <- ct.count[as.character(median_log10P$ct_top_auc)] %>% as.vector()

p <- ggscatter(median_log10P, x="n_cells", y="median.log10p",
               cor.coef = T, 
               cor.method = "spearman",
               xlab = "Number of cells per cell-type", 
               ylab = "Median observed link -log10(P)",
               label = "ct_top_auc",
               repel = T, cor.coef.coord = c(100,7))

ggsave(plot = p, "../figs/5_out/median-log10P.vs.cellcount.png", width = 7, height = 5)
```

## Scatter: Relation of celltype count vs max -Log10(P)

```{r}
Max.log10p <- links0.1 %>% 
  dplyr::group_by(ct_top_auc) %>% 
  summarise(Max.log10p = max(-log10(pvalue))) %>% 
  as.data.frame()

Max.log10p$n_cells <- ct.count[as.character(Max.log10p$ct_top_auc)] %>% as.vector()

p <- ggscatter(Max.log10p, x="n_cells", y="Max.log10p",
               cor.coef = T, 
               cor.method = "spearman",
               xlab = "Number of cells per cell-type", 
               ylab = "Max observed link -log10(P)",
               label = "ct_top_auc",
               repel = T, cor.coef.coord = c(100,135))

ggsave(plot = p, "../figs/5_out/max-log10P.vs.cellcount.png", width = 7, height = 5)
```

# Fig. 3 & S8-S9 - Prepare Epimap data
## Pseudobulk 

```{r}
# combine Epimap info with models tested
links0.1 <- readRDS("../results/links0.1_epimap.labels_ZINB.rds")
links0.1.multi <- readRDS("../results/links_signac_200null.r0.1.multimode.rds")
links0.1 <- merge(links0.1, 
                  links0.1.multi[,c("gene_peak", "H0.p", "multimodal")],
                  by = "gene_peak")

links0.1$Pearson_dist <- links0.1$score * exp(-links0.1$link.width/200000)
ct.epimap <- c("CD14_epimap","B_epimap","NK_epimap")

# include RegNMF
RegNMF <- readRDS("../results/RegNMF/RegNMF.scores_pbmc_links_hits.rds")
RegNMF$gene_peak <- gsub(":","-",RegNMF$gene_peak)
RegNMF <- RegNMF %>% dplyr::filter(gene_peak %in% links0.1$gene_peak)

RegNMF <- cbind(
  RegNMF,
  links0.1[match(RegNMF$gene_peak, links0.1$gene_peak), ct.epimap]
  )
Regnmf.cols <- c("CD14 Mono","B_cell","NK")
```

## By celltype subset
We need to get Epimap links overlap again because links might differ with pseudobulk links
```{r}
ct.ZINB <- readRDS("../results/celltype_subsetted.links0.1_ZINB.rds")

gr.ct.ZINB <- StringToGRanges(ct.ZINB$peak)

gr.ct.ZINB@elementMetadata <- ct.ZINB %>% 
  dplyr::select(-c(seqnames:strand)) %>% 
  DataFrame()

epimap.path <- "../data/Epimap/"

l <- lapply(ct.epimap, function(i){
  # subset by celltype
  epimap.gr <- readRDS(paste0(epimap.path, i, ".gr.rds"))
  sub.gr.ct.ZINB <- gr.ct.ZINB[gr.ct.ZINB$celltype == gsub("_epimap","",i)]
  
  # find overlap for subsetted PBMC test
  ovelaping_peaks <- findOverlaps(sub.gr.ct.ZINB, epimap.gr)
  
  df <- cbind(sub.gr.ct.ZINB@elementMetadata[ovelaping_peaks@from,], 
              epimap.gr@elementMetadata[ovelaping_peaks@to,]) %>% 
    as.data.frame()
  
  # keep overlapping peaks with same genes
  epimap.positive.links <- df %>% 
    filter(gene == gene_name) %>% 
    pull(gene_peak) %>% 
    unique()
  
  # label links validation TRUE if present in Epimap
  sub.gr.ct.ZINB@elementMetadata[[i]] <- ifelse(
    sub.gr.ct.ZINB$gene_peak %in% epimap.positive.links, 
    TRUE, 
    FALSE
  )
  
  df <- sub.gr.ct.ZINB@elementMetadata %>% as.data.frame()
  data.frame(gene_peak = df$gene_peak, 
             epimap.validated = df[[i]],
             celltype = df$celltype
  )
})

epimap.ct <- do.call(rbind,l)
ct.ZINB <- merge(ct.ZINB, epimap.ct, by=c("gene_peak","celltype"))
```

## Upset gene_peak overlap by Epimap celltype

```{r}
gr.links0.1 <- StringToGRanges(links0.1$peak)

gr.links0.1@elementMetadata <- DataFrame(links0.1)

epimap.path <- "../data/Epimap/"

lapply(ct.epimap, function(i){
  # subset epimap by celltype
  epimap.gr <- readRDS(paste0(epimap.path, i, ".gr.rds"))
  sub.gr.ct.ZINB <- gr.ct.ZINB[gr.ct.ZINB$celltype == gsub("_epimap","",i)]
  
  #####  find overlap for subsetted PBMC test
  ovelaping_peaks <- findOverlaps(sub.gr.ct.ZINB, epimap.gr)
  
  keep <- which(epimap.gr[ovelaping_peaks@to]$gene_name ==
                  sub.gr.ct.ZINB[ovelaping_peaks@from]$gene)
  ovelaping_links <- ovelaping_peaks[keep]
  
  # change Epimap gene_peak to match its overlaping peak in PBMC
  epimap.gr[ovelaping_links@to]$gene_peak <- sub.gr.ct.ZINB$gene_peak[ovelaping_links@from]

  ##### find overlap for pseudobulk PBMC test
  ovelaping_peaks <- findOverlaps(gr.links0.1, epimap.gr)
  
  keep <- which(epimap.gr[ovelaping_peaks@to]$gene_name ==
                  gr.links0.1[ovelaping_peaks@from]$gene)
  ovelaping_links <- ovelaping_peaks[keep]
  
  # change Epimap gene_peak to match its overlaping peak in PBMC
  epimap.gr[ovelaping_links@to]$gene_peak <- gr.links0.1$gene_peak[ovelaping_links@from]

  # labels
  ct <- gsub("_epimap","",i)
  upset_title <- paste0(ct, " links overlap")
  
  l_upset <- list(epimap.gr$gene_peak %>% unique(),
            sub.gr.ct.ZINB$gene_peak %>% unique(),
            links0.1$gene_peak %>% unique())
  names(l_upset) <- c(paste0("Epimap ", ct ," predictions"), 
                paste0("Subsetted ", ct ," links"), 
                "Pseudo-bulk PBMC links")
  
  # save plot
  png(paste0("../figs/5_out/upset.Epimap_",ct,".png"), res = 600, width = 6, height = 4, units = "in")
  print(upset(fromList(l_upset), order.by = "freq"))
  dev.off()
})
```

# Fig. 3 & S8-S9 - plot ROC Epimap 
## Pseudobulk
### Pearson R > 0.1

```{r}
tests <- list(c("zscore","Z-scores"),
              c("ZINB_Zvalue", "ZINB"),
              c("score","Pearson R"),
              c("Pearson_dist", "Pearson R * wDist"),
              c("dist", "1/Distance"))

links0.1$dist <- 1/links0.1$link.width

l.p <- lapply(1:3, function(i){
  l.df <- lapply(tests, function(test){
    
    links0.1.sub <- links0.1[!is.na(links0.1[[test[1]]]),]
    
    pred_ROCR <- prediction(abs(links0.1.sub[[test[1]]]), links0.1.sub[[ct.epimap[i]]])
    roc_ROCR <- performance(pred_ROCR, measure = "tpr", x.measure = "fpr")
    auc_ROCR <- performance(pred_ROCR, measure = "auc")
    
    data.frame(FPR = roc_ROCR@x.values[[1]], 
               TPR = roc_ROCR@y.values[[1]], 
               test = test[2],
               AUC = auc_ROCR@y.values[[1]])
  })
  df <- do.call(rbind,l.df)
  
  # RegNMF
  RegNMF.sub <- RegNMF[!is.na(RegNMF[[Regnmf.cols[i]]]),]
  
  pred_ROCR <- prediction(RegNMF.sub[[Regnmf.cols[i]]], RegNMF.sub[[ct.epimap[i]]])
  roc_ROCR <- performance(pred_ROCR, measure = "tpr", x.measure = "fpr")
  auc_ROCR <- performance(pred_ROCR, measure = "auc")
  
  df0 <- data.frame(FPR = roc_ROCR@x.values[[1]], 
                    TPR = roc_ROCR@y.values[[1]], 
                    test = "scREG",
                    AUC = auc_ROCR@y.values[[1]])  
  df <- rbind(df,df0)
  
  AUCs <- df %>% 
    group_by(AUC,test) %>% 
    summarise() %>% 
    arrange(desc(AUC)) %>% 
    as.data.frame()
  df$test <- factor(df$test, levels = AUCs$test)
  
  AUCs$AUC <- AUCs$AUC %>% round(3)
  
  AUCs.labs <- c()
  for (j in 1:nrow(AUCs)) {
    AUCs.labs <- c(AUCs.labs, paste0(AUCs[j,"test"], " = ", AUCs[j,"AUC"]))
  }
  
  ggplot(df, aes(FPR ,TPR, colour = test)) +
    geom_line() +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
    theme(legend.position = c(0.55, 0.3)) + 
    scale_color_discrete(labels = AUCs.labs) + 
    labs(colour = "AUC")
})

ggsave(plot = l.p[[1]], "../figs/5_out/pseudobulk/CD14_ROC.png", width = 8, height = 8)
ggsave(plot = l.p[[2]], "../figs/5_out/pseudobulk/B_ROC.png", width = 8, height = 8)
ggsave(plot = l.p[[3]], "../figs/5_out/pseudobulk/NK_ROC.png", width = 8, height = 8)
```

### Pearson R > 0.01

```{r}
links0.01 <- readRDS("../results/links0.01_Epimap.labeled.rds")

ct.epimap <- c("CD14_epimap","B_epimap","NK_epimap")

tests <- list(c("zscore","Z-scores"),
              c("score","Pearson R"))

l.p <- lapply(1:3, function(i){
  l.df <- lapply(tests, function(test){

    pred_ROCR <- prediction(abs(links0.01[[test[1]]]), links0.01[[ct.epimap[i]]])
    roc_ROCR <- performance(pred_ROCR, measure = "tpr", x.measure = "fpr")
    auc_ROCR <- performance(pred_ROCR, measure = "auc")
    
    data.frame(FPR = roc_ROCR@x.values[[1]], 
               TPR = roc_ROCR@y.values[[1]], 
               test = test[2],
               AUC = auc_ROCR@y.values[[1]])
  })
  df <- do.call(rbind,l.df)
  
  AUCs <- df %>% 
    group_by(AUC,test) %>% 
    summarise() %>% 
    arrange(desc(AUC)) %>% 
    as.data.frame()
  df$test <- factor(df$test, levels = AUCs$test)
  
  AUCs$AUC <- AUCs$AUC %>% round(3)
  
  AUCs.labs <- c()
  for (j in 1:nrow(AUCs)) {
    AUCs.labs <- c(AUCs.labs, paste0(AUCs[j,"test"], " = ", AUCs[j,"AUC"]))
  }
  
  ggplot(df, aes(FPR ,TPR, colour = test)) +
    geom_line() +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
    theme(legend.position = c(0.55, 0.3)) + 
    scale_color_discrete(labels = AUCs.labs) + 
    labs(colour = "AUC")
})

ggsave(plot = l.p[[1]], "../figs/5_out/pseudobulk/R0.01_CD14_ROC.png", width = 8, height = 8)
ggsave(plot = l.p[[2]], "../figs/5_out/pseudobulk/R0.01_B_ROC.png", width = 8, height = 8)
ggsave(plot = l.p[[3]], "../figs/5_out/pseudobulk/R0.01_NK_ROC.png", width = 8, height = 8)
```

## Subsetted celltypes

```{r}
celltypes <- factor(c("CD14","B","NK"))
tests <- list(c("zscore","Z-scores"),
              c("ZINB_Zvalue", "ZINB"),
              c("score","Pearson R"))

l.p <- lapply(1:3, function(i){
  l.df <- lapply(tests, function(test) {
    
    sub.ct.ZINB <- ct.ZINB %>% dplyr::filter(celltype == celltypes[i])
    sub.ct.ZINB <- sub.ct.ZINB[!is.na(sub.ct.ZINB[[test[1]]]),]

    pred_ROCR <- prediction(abs(sub.ct.ZINB[[test[1]]]), sub.ct.ZINB[["epimap.validated"]])
    roc_ROCR <- performance(pred_ROCR, measure = "tpr", x.measure = "fpr")
    auc_ROCR <- performance(pred_ROCR, measure = "auc")
    
    data.frame(FPR = roc_ROCR@x.values[[1]], 
               TPR = roc_ROCR@y.values[[1]], 
               test = test[2],
               AUC = auc_ROCR@y.values[[1]])
  })
  
  
  df <- do.call(rbind,l.df)
  
  AUCs <- df %>% 
    group_by(AUC,test) %>% 
    summarise() %>% 
    arrange(desc(AUC)) %>% 
    as.data.frame()
  df$test <- factor(df$test, levels = AUCs$test)
  
  AUCs$AUC <- AUCs$AUC %>% round(3)
  
  AUCs.labs <- c()
  for (j in 1:nrow(AUCs)) {
    AUCs.labs <- c(AUCs.labs, paste0(AUCs[j,"test"], " = ", AUCs[j,"AUC"]))
  }
  
  ggplot(df, aes(FPR ,TPR, colour = test)) +
    geom_line() +
    geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
    theme(legend.position = c(0.55, 0.3)) + 
    scale_color_discrete(labels = AUCs.labs) + 
    labs(colour = "AUC")
})


ggsave(plot = l.p[[1]], "../figs/5_out/celltype_subset/CD14_ROC.png", width = 8, height = 8)
ggsave(plot = l.p[[2]], "../figs/5_out/celltype_subset/B_ROC.png", width = 8, height = 8)
ggsave(plot = l.p[[3]], "../figs/5_out/celltype_subset/NK_ROC.png", width = 8, height = 8)
```

# Fig. 4 - Prepare CRISPRi data

```{r}
CRISPRi.validation <- readRDS("../results/CRISPR.validation_celltype_subsetted_ZINB.rds")
PB_CRISPRi <- readRDS("../results/links0.01_CRISPR_overlap.rds") %>% 
  as.data.frame() %>% 
  # See 4.Alternative_models.rmd chunk 7 for filters applied here
  dplyr::filter(gene_peak %in% CRISPRi.validation$gene_peak) %>% 
  distinct(gene_peak, .keep_all = T)

CRISPRi.validation.all <- CRISPRi.validation %>% 
  dplyr::filter(celltype == "all") %>% 
  dplyr::select(gene_peak,ZINB_Zvalue)

PB_CRISPRi <- merge(PB_CRISPRi, CRISPRi.validation.all, by = "gene_peak")

# RegNMF
RegNMF <- readRDS("../results/RegNMF/RegNMF.scores_pbmc_links_hits.rds")
Regnmf.cols <- c("CD14 Mono","B_cell","NK")
RegNMF$gene_peak <- gsub(":","-",RegNMF$gene_peak)
RegNMF <- RegNMF[,c(Regnmf.cols,"gene_peak")]
Regnmf.cols <- paste0("RegNMF_score_",Regnmf.cols)
colnames(RegNMF)[1:3] <- Regnmf.cols

PB_CRISPRi <- merge(PB_CRISPRi, RegNMF, by = "gene_peak", all.x=T)
```

## ROC CRISPRi pseudobulk

```{r}
tests <- list(c("zscore","Z-scores"),
              c("ZINB_Zvalue", "ZINB"), 
              c("score","Pearson R"),
              c("Pearson_dist", "Pearson * wDist"),
              c("dist", "1/Distance"))

PB_CRISPRi$dist <- 1/PB_CRISPRi$link.width
PB_CRISPRi$Pearson_dist <- PB_CRISPRi$score * (exp(-PB_CRISPRi$link.width/200000))

l.df <- lapply(tests, function(test){
  
  PB_CRISPRi.sub <- PB_CRISPRi[!is.na(PB_CRISPRi[[test[1]]]),]
  
  pred_ROCR <- prediction(abs(PB_CRISPRi.sub[[test[1]]]), PB_CRISPRi.sub[["Significant"]])
  roc_ROCR <- performance(pred_ROCR, measure = "tpr", x.measure = "fpr")
  auc_ROCR <- performance(pred_ROCR, measure = "auc")
  
  data.frame(FPR = roc_ROCR@x.values[[1]], 
             TPR = roc_ROCR@y.values[[1]], 
             test = test[2],
             AUC = auc_ROCR@y.values[[1]])
})

df <- do.call(rbind,l.df)


l.df <- lapply(1:3, function(i){
  
  PB_CRISPRi.sub <- PB_CRISPRi[!is.na(PB_CRISPRi[[Regnmf.cols[i]]]),]
  
  pred_ROCR <- prediction(PB_CRISPRi.sub[[Regnmf.cols[i]]], PB_CRISPRi.sub[["Significant"]])
  roc_ROCR <- performance(pred_ROCR, measure = "tpr", x.measure = "fpr")
  auc_ROCR <- performance(pred_ROCR, measure = "auc")
  
  data.frame(FPR = roc_ROCR@x.values[[1]], 
             TPR = roc_ROCR@y.values[[1]], 
             test = Regnmf.cols[i],
             AUC = auc_ROCR@y.values[[1]])  
})

df2 <- do.call(rbind,l.df)
df <- rbind(df,df2)
AUCs <- df %>% 
  group_by(AUC,test) %>% 
  summarise() %>% 
  arrange(desc(AUC)) %>% 
  as.data.frame()
df$test <- factor(df$test, levels = AUCs$test)

AUCs$AUC <- AUCs$AUC %>% round(3)

AUCs.labs <- c()
for (i in 1:nrow(AUCs)) {
  AUCs.labs <- c(AUCs.labs, paste0(AUCs[i,"test"], " = ", AUCs[i,"AUC"]))
}

AUCs.labs <- gsub("RegNMF_score_","scREG ", AUCs.labs)
AUCs.labs <- gsub("_"," ", AUCs.labs)
AUCs.labs <- gsub(" cell","", AUCs.labs)
AUCs.labs <- gsub(" Mono","", AUCs.labs)

cols <- c("#FF7856","#490092","#00F100","#999999","#FF50FF","#666666","#333333","#3FA0FF")
names(cols) <- c('Pearson * wDist', '1/Distance', 'Pearson R', 'RegNMF B_cell', 'ZINB', 'RegNMF_score_NK', 'RegNMF_score_CD14 Mono', 'Z-scores')

p <- ggplot(df, aes(FPR ,TPR, colour = test)) +
  geom_line() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  theme(legend.position = c(0.55, 0.3)) + 
  scale_color_manual(values = cols[AUCs$test], labels = AUCs.labs) + 
  labs(colour = "AUC") 

ggsave(plot = p, "../figs/5_out/pseudobulk/CRISPRi.png", width = 8, height = 8)
```

# Fig. S5 - Stat test association with a link's Cell count

```{r}
pbmc <- readRDS("../data/pbmc_multiome_processed.rds")

matg <- pbmc[["RNA"]]@counts[links0.1$gene,]
matp <- pbmc[["peaks2"]]@counts[links0.1$peak,]
matl <- matg * matp

links0.1$non0.cells.per.links <- apply(matl, 1, function(x) length(which(x>0)))
links0.1 <- merge(links0.1, RegNMF[,c("gene_peak","RegNMF_score_CD14 Mono")], by="gene_peak", all.x=T)
colnames(links0.1)[grep("CD14 Mono", colnames(links0.1))] <- "CD14.Mono_scREG"

# plot stats associations against non-zero cell counts
p1 <- ggplot(links0.1, aes(x = abs(zscore), y = non0.cells.per.links)) +
  geom_point(shape = ".") +
  scale_y_log10(limits = c(0.5, 11500)) +
  geom_smooth(color = "darkred", se = F, method = "lm") +
  ylab("Non-zero link cell count") +
  xlab("|Z-score|") +
  ggpubr::stat_cor(label.y = log10(11000)) 

p3 <- ggplot(links0.1, aes(x = ZINB_Zvalue, y = non0.cells.per.links)) +
  geom_point(shape = ".") +
  scale_y_log10(limits = c(0.5, 11500)) +
  xlim(0, 40) + # this removes 1 outlier
  geom_smooth(color = "darkred", se = F, method = "lm")+
  ylab("Non-zero link cell count")+
  xlab("|ZINB Z-value|") +
  ggpubr::stat_cor(label.y = log10(11000))

p4 <- ggplot(links0.1, aes(x = abs(score), y=non0.cells.per.links)) +
  geom_point(shape = ".") +
  geom_smooth(color = "darkred", se = F, method = "lm")+
  scale_y_log10(limits = c(0.5, 11500)) +
  ylab("Non-zero link cell count")+
  xlab("|Pearson R|") +
  ggpubr::stat_cor(label.y = log10(11000)) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank())

## add scREG 
p0 <- ggplot(links0.1, aes(x = CD14.Mono_scREG, y=non0.cells.per.links)) + 
  geom_point(shape = ".") +
  geom_smooth(color = "darkred", se = F, method = "lm")+
  scale_y_log10(limits = c(0.5, 11500)) +
  xlab(expression(paste(scREG[CD14]," score"))) +
  ggpubr::stat_cor(label.y = log10(11000)) +
  theme(axis.title.y = element_blank(), axis.text.y = element_blank()) +
  ylab("Non-zero link cell count")

patch2 <-  (p1 | p4 )/( p3 | p0)
ggsave(plot = patch2, "../figs/5_out/scatter.non_zero_cell_counts.vs.stats.png", width = 12, height = 12)
```

```{r}
sessionInfo()
```

